Supplementary Materials

### 2023-05-04 figures/fig-base-calling-error.tex
Merge suppl materials

\begin{document}
\begin{tikzpicture}[node distance=25mm, auto,
	dottedline/.style = {ultra thick, loosely dotted,shorten >=1mm, shorten <=1mm}]
\node (titlea) {\textbf{Base Calling Error FST}};

\node[block,below=2cm of titlea,fill=colorG!50] (start_sub) {start};
\node[block,right=15mm of start_sub] (S_sub) {S};
\node[block,above right=10mm and 30mm of S_sub] (ij_sub) {};
\node[block,right=30mm of S_sub] (ii_sub) {};
\node[block,below right=10mm and 30mm of S_sub] (in_sub) {};
\node[block,right=60mm of S_sub] (M_sub) {M};
\node[block,fill=colorR!50,right=15mm of M_sub] (end_sub) {end};

@@

\draw[line] (M_sub) to[bend right=20] node[e] {i\textrightarrow{}j} (ij_sub);
\draw[line] (ij_sub) to[bend right=20] node[p] {$\varepsilon$} (S_sub);
\draw[line] (M_sub) to node[e] {i\textrightarrow{}i} (ii_sub);
\draw[line] (ii_sub) to node[p] {$1 - 3 \cdot \varepsilon$} (S_sub);
\draw[line] (M_sub) to[bend right=-20] node[e] {i\textrightarrow{}N} (in_sub);
\draw[line] (in_sub) to[bend right=-20] (S_sub);

@@

\node[below=2cm of start_sub,text width=50mm,anchor=north west] (sequences) {
\begin{tabular}{r@{\,: }l}
\multicolumn{2}{l}{\textbf{Sequences}}\\
    \hline
	i & input nucleotides\\
	j & intermediate nucleotides\\
	N & input ambiguous nucleotide N\\
\end{tabular}
};

\node[above right=-1.03cm and 5mm of sequences,text width=50mm] {
\begin{tabular}{r@{\,: }l}
\multicolumn{2}{l}{\textbf{Parameters}}\\
\hline
	$\varepsilon$ & base calling error weight\\
\end{tabular}
};

>>>>

\begin{document}
\begin{tikzpicture}[node distance=25mm, auto,
	dottedline/.style = {ultra thick, loosely dotted,shorten >=1mm, shorten <=1mm}]
%\node (titlea) {\textbf{Base Calling Error FST}};

\node[block,fill=colorG!50] (start_sub) {start};
\node[block,right=15mm of start_sub] (S_sub) {S};
\node[block,above right=10mm and 30mm of S_sub] (ij_sub) {};
\node[block,right=30mm of S_sub] (ii_sub) {};
\node[block,below right=10mm and 30mm of S_sub] (in_sub) {};
\node[block,right=60mm of S_sub] (M_sub) {M};
\node[block,fill=colorR!50,right=15mm of M_sub] (end_sub) {end};

@@

\draw[line] (M_sub) to[bend right=20] node[e] {i\textrightarrow{}j} (ij_sub);
\draw[line] (ij_sub) to[bend right=20] node[w] {$\varepsilon$} (S_sub);
\draw[line] (M_sub) to node[e] {i\textrightarrow{}i} (ii_sub);
\draw[line] (ii_sub) to node[w] {$1 - 3 \cdot \varepsilon$} (S_sub);
\draw[line] (M_sub) to[bend right=-20] node[e] {i\textrightarrow{}N} (in_sub);
\draw[line] (in_sub) to[bend right=-20] (S_sub);

@@

\node[below right=2cm and -1.75cm of start_sub,text width=50mm,anchor=north west] (sequences) {
\begin{tabular}{r@{\,: }l}
\multicolumn{2}{l}{\textbf{Sequences}}\\
    \hline
	$i \rightarrow i$ & matching input to intermediate base\\
	$i \rightarrow j$ & mismatching input to intermediate base\\
	$i \rightarrow N$ & input to intermediate ambiguous base N\\
\end{tabular}
};

\node[above right=-1.03cm and 25mm of sequences,text width=50mm] {
\begin{tabular}{r@{\,: }l}
\multicolumn{2}{l}{\textbf{Parameters}}\\
\hline
	$\varepsilon$ & base calling error weight\\
\end{tabular}
};

### 2023-05-04 results_discussion.tex
Merge suppl materials

After downloading, we removed 2232 sequences longer than 6000 nucleotides, identified 8369 sequence pairs that contained gaps identified by at least one aligner, and 5399 ungapped sequence pairs.
We then randomly introduced gap patterns extracted from all five methods into the ungapped sequence pairs to generate the benchmark alignments.
Alignment accuracy was measured using the distance metric $d_{seq}$
\citembe{metrics_blackburne_whelan_2011} between simulated and inferred
alignments.
In addition, accuracy of positive and negative selection was calculated
using the $F_1$ score by estimating $k_s$ and $k_a$ statistics
\citembe{ka_ks_li_1993}.
The $F_1$ score evaluates the accuracy of a model by assigning equal importance to precision and recall, and ranges between 0 and 1, with a score of 1 representing a perfect result.

% Software comparison table
\begin{table}[!ht]
\centering
\input{figures/table-comp}
 \caption{COATi generates better alignments than other alignment algorithms. Results of COATi, PRANK, MAFFT, Clustal$\Omega$, and MACSE aligning 5399 empirically simulated sequence pairs. Perfect alignments have $d_{seq}=0$, best alignments have the lowest $d_{seq}$, and imperfect alignments have $d_{seq}>0$ when at least one aligner found a perfect alignment.}
 \label{table:comp}
\end{table}

>>>>

After downloading, we removed 2232 sequences longer than 6000 nucleotides, identified 6048 sequence pairs that contained gaps identified by at least one aligner, and 7719 ungapped sequence pairs.
We then randomly introduced gap patterns extracted from all five methods into the ungapped sequence pairs to generate the benchmark alignments.
Alignment accuracy was measured using the distance metric $d_{seq}$
\citembe{metrics_blackburne_whelan_2011} between simulated and inferred
alignments.
In addition, accuracy of positive and negative selection was calculated
using the $F_1$ score by estimating $k_s$ and $k_a$ statistics
\citembe{ka_ks_li_1993}.
The $F_1$ score evaluates the accuracy of a model by assigning equal importance to precision and recall, and ranges between 0 and 1, with a score of 1 representing a perfect result.

% Software comparison table
\begin{table}[!ht]
\centering
\input{figures/table-comp}
 \caption{COATi generates better alignments than other alignment algorithms. Results of COATi, PRANK, MAFFT, Clustal$\Omega$, and MACSE aligning 7719 empirically simulated sequence pairs. Perfect alignments have $d_{seq}=0$, best alignments have the lowest $d_{seq}$, and imperfect alignments have $d_{seq}>0$ when at least one aligner found a perfect alignment.}
 \label{table:comp}
\end{table}
